Probabilistic Policy Violation Detection with IPFWD

### 2017-02-06 paper/main.tex
Wrote IPFWD section, fix graphs

\author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
\thanks{Dell EMC Isilon}}

@@

% What does it do?
IPFWD increases firewall performance by reducing the average amount of work it
takes to process a packet. Depending on the current system load, IPFWD
increases or decreases the chance the IPFW will immediately accept a given
packet. It does this by claiming and maintaining an early rule in the IPFW rule
set of the following form: 'prob 0.000 allow ip from any to any'.

% How does it work?
% - Probabilistic packet matching
% - Log extension

% Explain rationale for it acting as it does

% When should this system be used?
% - High performance
% - Resource conscious
\blindtext

@@

% Talk about shift in mindset from "blocking attacks" to "detecting attacks"
% - Why is this reasonable?
% - Is this really different than traditional firewalls?
\blindtext

@@

  \begin{center}
  \includegraphics[width=0.45\textwidth]{isilon40_cpu}
  \includegraphics[width=0.45\textwidth]{isilon10_cpu}
  \includegraphics[width=0.45\textwidth]{wwu1_cpu}
  \includegraphics[width=0.45\textwidth]{virtual10_cpu}
  \end{center}

@@

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

>>>>

\author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
\thanks{In collaboration with Dell EMC Isilon}}

@@

% What does it do?
IPFWD increases firewall performance by reducing the average amount of work it
takes to process a packet. Depending on the current system load, IPFWD
increases or decreases the chance the IPFW will immediately accept a given
packet. It does this by adding and maintaining an early rule in the IPFW rule
set of the following form: \verb|prob 0.000 allow ip from any to any|.

% How does it work?
% - Probabilistic packet matching
Since IPFW uses a first-match-wins rule system, this early rule will be
encountered before the majority of the other rules in the rule set.  The
probability given in this rule determines the chance that we skip processing
all further rules, supposing that there hasn't been a match already.

This is especially valuable for a whitelist firewall rule set dealing with
large amounts of rejected traffic; normally each rejected packet has all rules
applied to it while IPFW looks for a match.  When no matches are encountered,
the default rule is employed and the packet is rejected.

In a typical data center environment the vast majority of network traffic is
valid and going to be accepted eventually. However, these rule sets are often
complex and take increasing time to match a given packet depending on the
length. For a rule set of length $n$, we can expect to a packet to encounter
$n/2$ failed matches before matching the correct rule and being accepted.

IPFWD improves the average number of failed matches before successful match
from

\[
\frac{n}{2}
\]

to

\[
\Big(k \cdot \mathsf{P}(\textit{EA})\Big) + 
\Big(\frac{n}{2} \cdot \mathsf{P}(\neg{\textit{EA}})\Big) 
\]

where $k$ is number of rules before the early acceptance rule and
$\mathsf{P}(\textit{EA})$ is the probability that the early acceptance rule is
matched.

Removing the constant $k$, IPFWD reduces the number of failed matches before a
successful match by a factor of

\[
\frac
    {\frac{n}{2}}
    {\frac{n}{2} \cdot \mathsf{P}(\neg{\textit{EA}})} 
= \frac{1}{\mathsf{P}(\neg{\textit{EA})}}.
\]

% - Log extension
However, while this method increases the speed the system processes packets, it
creates a chance that packets that would have normally been rejected will get
through the firewall. We'll refer to these types of packets as invald.

To account for this, IPFWD extends the IPFW system logs to include information
about the early accpetance probability. From this information, administrators
can extrapolate the probability that additional invalid packets made it through
the firewall for each that is detected. This is discussed further in the
Security Implications section.

% When should this system be used?
IPFWD is not intended to be used on all types of systems. It's application
instead most directly benefits the following types of systems:

\begin{enumerate}
    \item \textbf{High Network Performance}\\
      The gains provided by IPFWD are only apparent in systems were network
      throughput is matched by or out performs CPU speed. Systems that can
      already easily handle fully processing each packet will not benefit by
      reducing processing time. Given modern CPU clock speeds, systems with
      network cards slower than 1 Gb/sec will likely not benefit either.

    \item \textbf{Resource Conscious}\\
      In line with high performance systems, systems with strict resource
      restrictions can benefit by the reduced network related CPU usage.

    \item \textbf{Complex Firewall Requirements}\\
      IPFWD works by reducing the average number of rules that fail to match
      before a packet is finished being processed. Very simple rulesets will
      gain little from reducing the number of rules applied to each packet.

    \item \textbf{Detection Based Security Policy}\\ 
      By using IPFWD, administrators sacrifice the assurance that every invalid
      packet will be rejected.  However, they are still provided information as
      to whether some invalid packets were rejected, and that information can
      be used to infer the existance of additional invalid packets.

\end{enumerate}

% Explain rationale for it acting as it does
IPFWD is based on the assumptions that the majority of network traffic is
valid, malicious events are rare, and

@@

% Talk about shift in mindset from "blocking attacks" to "detecting attacks"
% - Why is this reasonable?
% - Is this really different than traditional firewalls?

@@

  \begin{center}
  \includegraphics[width=0.45\textwidth]{cpu_isilon40}
  \includegraphics[width=0.45\textwidth]{cpu_isilon10}
  \includegraphics[width=0.45\textwidth]{cpu_wwu1}
  \includegraphics[width=0.45\textwidth]{cpu_virtual10}
  \end{center}

@@
